Big Ideas in Sports Analytics and Statistical Tools for their Investigation

### 2022-11-15 sports_analytics.Rmd
framing for sports betting; #29

# Sports betting

@@

## Betting pools

>>>>

# Non-performance data

Most of the research in sports analytics is fueled by the analysis of data recorded from the outcome of sports contests.
However, a growing body of literature is informed by non-performance data about sports. 
This research includes fields like sports economics and kinesiology that are subfields of mature disciplines (a treatment of either is well beyond the scope of the this paper).
In addition to this body of research, we note that today, professional sports teams are collecting biometric data on their own players, and presumably using that data to learn about how their workouts, sleep patterns, and diets impact their athletic performance. 
While this research may constitute "emerging methodologies," it will likely never be published, because each team will use it to their competitive advantage.

Nevertheless, in this Section we discuss concepts stemming from one notable, relatively new source of (semi-)public non-performance sports data: data from betting markets.
Since the 2018 United States Supreme Court decision in [*Murphy v. National Collegiate Athletic Association*](https://en.wikipedia.org/wiki/Murphy_v._National_Collegiate_Athletic_Association), sports gambling has exploded in the U.S. 
The increasing interest in sports gambling has led to increasing interest in sports gambling *data*, and that data has proven useful to researchers in two major ways.

First, betting market data is probably the best source for estimating the true probability of a team winning a game. 
The efficiency of betting market data in this respect has been demonstrated time and time again.
The utility of these estimates have then informed research that has helped us learn about the sports themselves.
In this sense, data generated by sports gambling has been an important source of data useful for sports analytics (see Section \ref{sec:betting-market-data}).

Second, sports analytics researchers have studied various types of sports gambling outlets (including fantasy sports). 
This research has estimated probabilities, evaluated common strategies, and offered optimized approaches for a variety of different games of chance (see Section \ref{sec:pools}).

## Sports betting market data {#sec:betting-market-data}

@@

## Analytics for sports betting pools {#sec:pools}
